### 2014-04-08 paper/relatedwork.tex
bjoern's pass through intro, related work (almost complete)

\section{Related Work}\tovi{related work section still needs some work}
Our contributions relate to other efforts aiming to add interactivity to digitally fabricated objects; and to techniques for routing and model modification.

@@

Previous work has investigated the integration of interactivity with the fabrication process.  Printed Optics (\cite{Willis-printedoptics}) used light pipes, solid clear tubes, to create integrated displays and touch sensors. While the research describes a process, it does not provide a design tool for creating such objects. Our work is inspired by Printed Optics and introduces a design tool for modelling objects that can have these types of capabilities.

Slyper, et al., 3D printed flexible robot armatures capable of sensing interaction via changes in air pressure \cite{Slyper-pressure}.  They created all their models by hand, and were forced to place interaction locations in areas accessible to electronics assembly: this technique is complementary to ours, as we allow placing interactive elements anywhere on an object's surface. Sauron investigated computer vision to track interactions with physical mechanisms \cite{Savage-sauron}.  While Savage, et al., used mirrors for redirection, we believe that the required line-of-sight operation for their computer vision techniques may be possible using printed optics or inserted optical fibers \george{huh?}.

General-purpose routing for connecting 2D capacitive sensors to terminals is discussed in Midas \cite{Savage-midas}.  This technique, based on circuitboard routing, is limited to planar and unfoldable shapes, and intended for use on the surface of objects rather than through their cores.  Fabrickation works towards rapid fabrication of functional objects by integrating construction kit building blocks \cite{Mueller-fabrickation}.  While we recognize that 3D printing requires significant time, we believe creating precise internal structures in 3D printed objects is a strong case for digital fabrication over manual or traditional methods. \george{wording}

\subsubsection{Fabrication + Electronics}

Other research has investigated the integration of traditional electronics into 3D prints by depositing non-printable materials after printing.  Sells, et al., \cite{Sells-reprap}, describe using a heated syringe to deposit metal in printed 2D circuitboard channels on objects' surfaces.  Sarik, et al., \cite{Sarik-tracebrush} and the Optomec company \cite{optomec} describe \emph{spraying} conductive material onto printed objects to create circuitry.  Injection of liquid metals for sensing in softer-than-skin electronics was explored by Park, et al., \cite{Park-microchannels} and Majidi, et al., \cite{Majidi-curvature}.  Navarette, et al., created a functioning GPS routed in true 3D (not multilayer boards' 2.5D), hand-built using a one-off routing.  Our work differs from these because we offer \emph{algorithms and techniques} for \emph{internal}, \emph{true 3D} routing.

\subsubsection{Interaction for Fabrication}

Other researchers are investigating interaction for fabrication (i.e., interaction techniques for controlling digital fabrication machines) rather than fabrication for interaction as we do.  This includes work like Constructable \cite{Mueller-constructable}, KidCAD \cite{Follmer-kidcad}, and Willis, et al,'s interactive interfaces \cite{Willis-interactive}. Their goals and contributions are orthogonal to ours.

@@

Similar to our path routing algorithm, Wong presents a graph-based approach for generating continuous line illustrations from images \cite{Wong-continuousline}. Its approach of linking edges into a continuous line based on semi-Eulerization is similar to ours.  However, we are not confined to the plane to avoid line-crossings, and we can create ``invisible'' lines using \emph{blocking out}\footnote{In neon sign making, segments of a tube that should not appear illuminated are covered with opaque paint. This is called "blocking out". \url{http://en.wikipedia.org/wiki/Neon_sign}} \george{too much detail and talk of our work}. Prior work used tours of nodes, rather than edges, and cast continuous line drawing as a traveling salesman problem~\cite{Bosch-tsp}. In traditional neon sign making, ``routing'' is a conversation between a graphic designer and a glass bender on what is feasible and comfortable for the latter and what will help him avoid burning his hands and wasting time or glass \cite{strattman1997neon}.

@@

>>>>

%!TEX root = series of tubes.tex
\section{Related Work}\tovi{related work section still needs some work}
Our contributions relate to other efforts aiming to add interactivity to digitally fabricated objects; and to techniques for routing and model modification.

@@

Previous work has begun to investigate how to fabricate interactive objects using particular sensing and display technologies. Printed Optics (\cite{Willis-printedoptics}) uses light pipes to create integrated displays and sensed input optically through components added at print time; while Slyper, et al., 3D print flexible robot armatures capable of sensing interaction via changes in air pressure \cite{Slyper-pressure}. Our work is inspired by these projects, and contributes a general design tool for modelling objects that can have such capabilities.

Sauron tracks interactions with printed physical mechanisms by placing a camera inside hollow objects, where all components have to be visible to a camera~\cite{Savage-sauron}.  Our technique instead can work with general, solid models and only carves out pipes for interaction where needed. Midas generates fabrication files custom capacitive touch sensors~\cite{Savage-midas}.  This technique, based on circuitboard routing, is limited to planar and unfoldable shapes. Connections are routed on the surface of objects. \systemname can also be used for touch sensing, but routes conducters through the inside of objects.

Fabrickation works towards rapid fabrication of functional objects by integrating construction kit building blocks \cite{Mueller-fabrickation}.  While we recognize that 3D printing requires significant time, we believe creating precise internal structures in 3D printed objects is a strong case for digital fabrication over manual or traditional methods. \george{wording}\bjoern{needs work}

\subsubsection{Fabrication with Printed Electronics}
Researchers are also investigating how to print conductors and place electronic components as part of additive manufacturing.  Sells, et al., \cite{Sells-reprap}use a heated syringe to deposit metal in printed 2D circuitboard channels on objects' surfaces.  Sarik, et al., \cite{Sarik-tracebrush} use aerosol jetting of conductive material onto printed objects.  Park, et al., \cite{Park-microchannels} and Majidi, et al., \cite{Majidi-curvature} inject liquid metals. Our approach does not deposit such materials, but creates structures inside a 3D model that make manual post-print injection of conductive ink possible. 
%Navarette, et al., created a functioning GPS routed in true 3D (not multilayer boards' 2.5D), hand-built using a one-off routing.  \bjoern{need more context for this}
%Our work differs from these because we offer \emph{algorithms and techniques} for \emph{internal}, \emph{true 3D} routing.

@@

Similar to our path routing algorithm, Wong presents a graph-based approach for generating continuous line illustrations from images \cite{Wong-continuousline}. Its approach of linking edges into a continuous line based on semi-Eulerization is similar to ours.  However, we are not confined to the plane to avoid line-crossings, and we can create ``invisible'' lines using \emph{blocking out}\footnote{In neon sign making, segments of a tube that should not appear illuminated are covered with opaque paint. This is called "blocking out". \url{http://en.wikipedia.org/wiki/Neon_sign}} \george{too much detail and talk of our work}. Prior work used tours of nodes, rather than edges, and cast continuous line drawing as a traveling salesman problem \cite{Bosch-tsp}. In traditional neon sign making, ``routing'' is a conversation between a graphic designer and a glass bender on what is feasible and comfortable for the latter and what will help him avoid burning his hands and wasting time or glass \cite{strattman1997neon}.

@@

\subsubsection{Interaction for Fabrication vs. Fabrication for Interaction}
HCI researchers are also developing novel interaction techniques for providing input to digital fabrication machines~\cite{Mueller-constructable,Willis-interactive,mixfab}, and new hybrid hand tools combine manual processes with digital fabrication~\cite{zoran2013freed,rivers2012position}. These projects are complementary as they target different aspects of the design process, though it is conceivable that our pipe design tools could be added to augmented reality modeling interfaces like MixFab~\cite{mixfab}.
